Population structure along the genome

### 2016-01-29 writeup/local_pca_paper.tex
added more refs

Review of kinship matrix: 
it's either an expected kinship, given the pedigree;
or an estimated genome-wide average.
Wright's path coefficients \citep{wright1943isolation}.
Why it helps with confounding for GWAS.
Graham \& Vince's paper, maybe.
Like IBD, is only well-defined in a known pedigree, up to the founders.

@@

Review of selection causing differential patterns along the genome.
Locally everything is treelike (gene trees);
kinship matrix is an average of these (write equation for this).
Selective sweeps cause local recent ancestry/short trees.
Balancing selection causes deep trees.
Background selection, shallow ones.
Refs: hitchhiking \citep{maynardsmith1974hitchhiking},
selection on LD \citep{mcvean2007structure},

@@

What is "population structure"?
Asks which "populations" are closely related, more diverged, how much diversity do they harbor.
Often geographic.
Vital in exploratory data analysis.
It is a summary of kinship: 
lack of migration between pops causes a deficit in connections through the pedigree,
and so affects kinship.

Review of methods for visualizing pop structure:
PCA, etcetera.
Include maps of homozygosity.

Other semi-related stuff:
estimation of covariance matrices;
local pca(?);

@@

humans:
effects of selection \citep{sabeti2006positive}
dataset(s)

>>>>

Review of kinship matrix: 
it's either an expected kinship, given the pedigree;
or an estimated genome-wide average.
Wright's path coefficients \citep{wright1943isolation}.
Why it helps with confounding for GWAS.
Graham \& Vince's paper, maybe.
Like IBD, is only well-defined in a known pedigree, up to the founders.
Kinship and confounding reviewed in \citet{astle2009population}.

@@

Review of selection causing differential patterns along the genome.
Locally everything is treelike (gene trees);
kinship matrix is an average of these (write equation for this).
Selective sweeps cause local recent ancestry/short trees.
Balancing selection causes deep trees.
Background selection, shallow ones.
Extreme examples of free gene flow in some places between species: e.g. Heliconius.
Introgression may be nonuniform, e.g. neanderthal, others.
Refs: hitchhiking \citep{maynardsmith1974hitchhiking},
\citet{kim2011hitchhiking} study hitchhiking in a spatially subdivided population,
\citet{mcvean2007structure} looks at the effect of selection on LD.
\citet{barton2000genetic} reviews hitchhiking.
\citet{bierne2010distinctive} discusses how hitchhiking effect decreases with geographic distance.
\citet{charlesworth2003review} reviews patterns of diversity, relating spatial structure to effects of selection.

@@

What is "population structure"?
Asks which "populations" are closely related, more diverged, how much diversity do they harbor.
Often geographic.
Vital in exploratory data analysis.
It is a summary of kinship: 
lack of migration between pops causes a deficit in connections through the pedigree,
and so affects kinship.
Wright defined $F_{ST}$ in \citep{wright1949genetical}, and says ``It has probably occurred to the reader that the coefficientof inbreeding may mean very different
things in different cases.''

Review of methods for visualizing pop structure:
PCA, structure \citep{falush2003inference}, EEMS, \citep{petkova2014visualizing}, \citep{yang2012modelbased},
Maps of heterozygosity \citep{ramachandran2005support}.
Genealogical interpretation of PCA by \citet{mcvean2009genealogical}.
Other semi-related stuff:
estimation of covariance matrices;
local pca(?);

@@

We apply this to three species.

humans:
effects of selection are seen \citep{sabeti2006positive}.
\citet{duforetfrebourg2015detecting} showed that variants known to underly locally adaptive traits correlated most strongly with PCs.
\citet{tian2008analysis} noted that removing the segregating inversion on chromosome 8 changed the PCA for a set of 1400 Europeans
dataset(s)
